### 2016-01-20 paper.tex
for --> of
Tense consistency; resolves #50

To calculate turbine RPM from shaft angle, the encoder signal is differentiated
using a second order central difference scheme, after which an 8 sample wide
moving average smoothing filter is applied to match the noise level present in
the redundant turbine RPM measurement from the motion controller. A similar
approach is used for calculating tow carriage speed $U_\infty$ from carriage
position measurements. Power and drag coefficients are calculated as
instantaneous quantities from the carriage speed as

@@

Mean rotor power coefficients for multiple Reynolds numbers are plotted versus
tip speed ratio in Fig.~\ref{fig:cp-curves}. $C_P$ is shown to increase in
general with $Re$, along with a reduction in the optimal tip speed ratio, due to
the tendency of foils to stall at higher angles of attack at higher $Re$, and
the higher angle of attack ranges seen by the blades at lower $\lambda$. These
effects diminish with increasing $Re$, which is expected as the blade boundary
layers transition to turbulence closer to the leading edge \cite{Lissaman1983,
    McMasters1980, Bachant2015-RVAT-Re-dep}, which helps flow remain attached longer
as it moves against the adverse pressure gradient on the suction side of the
foil. Note how at lower tip speed ratios and Reynolds numbers, $C_P$ becomes
significantly negative, which is possible thanks to the speed control of the
experimental setup's servo motor applying negative torque to the rotor.

@@

Strut drag losses do not change much for the streamlined NACA 0021 struts in the
towed versus stationary configuration, which helps explain why overall rotor
drag coefficients remain relatively constant even at low Reynolds number, were
$C_P$ is dramatically reduced, or even negative. For the cylindrical struts,
losses increase significantly when towed and operating in the mid range of tip
speed ratios. Though a real turbine would never use such high drag struts, with
respect to numerical modeling, these measurements will allow support struts to
be simulated without blades, to isolate and evaluate the ability to predict
these losses independent of the blade loading.

@@

Derivatives were computed with a second order central difference scheme for
interior points, and a second order inward-facing scheme for the edges,
identical to \cite{Bachant2015-JoT}. Weighted averages for these computation are
shown in Fig.~\ref{fig:Ktransport}. Due to the weaker blade vortex shedding,
transport due to vertical advection at this point in the wake is approximately 3
times lower than the higher solidity UNH-RVAT. Note that direct comparison is
somewhat invalid, since the total measurement plane area is about 5\% lower for
this experiment. However, the differences observed are beyond this discrepancy.
We also see relatively lower levels of cross-stream turbulent transport due to
the lack of blade stall vortex shedding. These results may have interesting
implications regarding the application of turbines with lower power coefficient
to possibly improve overall array performance through enhanced transport of
kinetic energy from the free stream, though evaluating these trade-offs will
require a detailed analysis of the downstream evolution and turbine--wake
interaction.

@@

The performance and near-wake for a 1:6 scale DOE Reference Model 2 cross-flow
turbine was measured in a towing tank. Performance was assessed for Reynolds
number dependence, showing a convergence to a weakly $Re$-dependent linear
regime at approximately $Re_D \approx 1 \times 10^6$ or $Re_{c,\mathrm{ave}}
\approx 2 \times 10^5$. Comparison was made between this turbine and the higher
solidity UNH-RVAT, tested in nearly identical conditions, which showed similar
Reynold number thresholds but a flatter linear regime, likely to due virtual
camber of the higher chord-to-radius ratio of the UNH-RVAT blades.

The wake at $x/D=0.93$ downstream was shown to be relatively symmetrical and
lacked the evidence of strong blade stall, which once again differentiates this
turbine from the UNH-RVAT. Terms from the mean kinetic energy transport equation
were also computed in this $y$--$z$ plane, showing the relative importance of
the vertical advection compared with turbulent transport terms at this location.
It is interesting to note that though the turbine is a more effective energy
converter than the UNH-RVAT, its wake recovery may in fact be slower due to
weaker blade tip vortex shedding.

The effects of parasitic drag from blade support struts on turbine performance
were measured by rotating the turbine without blades while stationary and while
towing. It is shown that these losses---even for a streamlined hydrofoil
strut---can become significant quickly at higher tip speed ratios---up to an
approximate 5 percentage point decrease in power coefficient at a tip speed
ratio of 5. These measurements were repeated with a set of high-drag cylindrical
struts, which as expected, prevented the turbine from producing any mechanical
power at any tip speed ratio. Nevertheless these measurements provide important
validation data for both high and low fidelity numerical performance prediction
models, allowing researchers to test models without blade effects.

>>>>

To calculate turbine RPM from shaft angle, the encoder signal was differentiated
using a second order central difference scheme, after which an 8 sample wide
moving average smoothing filter was applied to match the noise level present in
the redundant turbine RPM measurement from the motion controller. A similar
approach was used for calculating tow carriage speed $U_\infty$ from carriage
position measurements. Power and drag coefficients are calculated as
instantaneous quantities from the carriage speed as

@@

Mean rotor power coefficients for multiple Reynolds numbers are plotted versus
tip speed ratio in Fig.~\ref{fig:cp-curves}. In general, $C_P$ increases with
$Re$, along with a reduction in the optimal tip speed ratio, due to the tendency
of foils to stall at higher angles of attack at higher $Re$, and the higher
angle of attack ranges seen by the blades at lower $\lambda$. These effects
diminish with increasing $Re$, which is expected as the blade boundary layers
transition to turbulence closer to the leading edge \cite{Lissaman1983,
    McMasters1980, Bachant2015-RVAT-Re-dep}, which helps flow remain attached longer
as it moves against the adverse pressure gradient on the suction side of the
foil. Note how at lower tip speed ratios and Reynolds numbers, $C_P$ becomes
significantly negative, which was possible thanks to the speed control of the
experimental setup's servo motor applying negative torque to the rotor.

@@

Strut drag losses did not change much for the streamlined NACA 0021 struts in
the towed versus stationary configuration, which helps explain why overall rotor
drag coefficients remained relatively constant even at low Reynolds number,
where $C_P$ was dramatically reduced, or even negative. For the cylindrical
struts, losses increased significantly when towed and operating in the mid range
of tip speed ratios. Though a real turbine would never use such high drag
struts, with respect to numerical modeling, these measurements will allow
support struts to be simulated without blades, to isolate and evaluate the
ability to predict these losses independent of the blade loading.

@@

Derivatives were computed with a second order central difference scheme for
interior points, and a second order inward-facing scheme for the edges,
identical to \cite{Bachant2015-JoT}. Weighted averages for these computation are
shown in Fig.~\ref{fig:Ktransport}. Due to the weaker blade vortex shedding,
transport due to vertical advection at this point in the wake was approximately
3 times lower than the higher solidity UNH-RVAT. Note that direct comparison is
somewhat invalid, since the total measurement plane area was about 5\% lower for
this experiment. However, the differences observed are beyond this discrepancy.
We also see relatively lower levels of cross-stream turbulent transport due to
the lack of blade stall vortex shedding. These results may have interesting
implications regarding the application of turbines with lower power coefficient
to possibly improve overall array performance through enhanced transport of
kinetic energy from the free stream, though evaluating these trade-offs will
require a detailed analysis of the downstream evolution and turbine--wake
interaction.

@@

The performance and near-wake of a 1:6 scale DOE Reference Model 2 cross-flow
turbine was measured in a towing tank. Performance was assessed for Reynolds
number dependence, showing a convergence to a weakly $Re$-dependent linear
regime at approximately $Re_D \approx 1 \times 10^6$ or $Re_{c,\mathrm{ave}}
\approx 2 \times 10^5$. Comparison was made between this turbine and the higher
solidity UNH-RVAT, tested in nearly identical conditions, which showed similar
Reynold number thresholds but a flatter linear regime, likely to due virtual
camber of the higher chord-to-radius ratio of the UNH-RVAT blades.

The wake at $x/D=0.93$ downstream was shown to be relatively symmetrical and
lacked the evidence of strong blade stall, which once again differentiates this
turbine from the UNH-RVAT. Terms from the mean kinetic energy transport equation
were also computed in this $y$--$z$ plane, showing the relative importance of
the vertical advection compared with turbulent transport terms at this location.
It is interesting to note that though the RM2 is a more effective energy
converter than the UNH-RVAT, its wake recovery may in fact be slower due to
weaker blade tip vortex shedding and lower levels of turbulence in the
near-wake.

The effects of parasitic drag from blade support struts on turbine performance
were measured by rotating the turbine without blades while stationary and while
towing. It is shown that these losses---even for a streamlined hydrofoil
strut---can become significant quickly at higher tip speed ratios---up to an
approximate 5 percentage point decrease in power coefficient at a tip speed
ratio of 5. These measurements were repeated with a set of high-drag cylindrical
struts, which as expected, prevented the turbine from producing any mechanical
power at any tip speed ratio. Nevertheless these measurements provide important
validation data for both high and low fidelity numerical performance prediction
models, allowing researchers and engineers to test predictions without blade
effects.
